Supplemental Information: Aligning marine species range data to better serve science and conservation

### 2016-09-02 case_study_mpa/3_mpa_spp_analysis.Rmd

print(barchart_mpa_gap)
ggsave(plot = barchart_mpa_gap,
       filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
       width = 11.4, height = 5, units = 'cm', dpi = 600)
# ggsave(plot = barchart_mpa_gap,
#        filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
#        width = 11.4, height = 5, units = 'cm', dpi = 600)

>>>>

print(barchart_mpa_gap)
ggsave(plot = barchart_mpa_gap,
       filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
       width = 19.05, height = 5, units = 'cm', dpi = 600)
# ggsave(plot = barchart_mpa_gap,
#        filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
#        width = 11.4, height = 5, units = 'cm', dpi = 600)

### 2016-09-02 clip_fao/fao_clip_pac.Rmd

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
                            layer = 'fao_rgns2')
rast_ext  <- extent(c(85, 220, -59, 50))
fao_polys <- crop(fao_polys, rast_ext)

World1 <- crop(wrld_simpl, rast_ext)

@@

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum
spp_cells_clipped <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_e$am_sid) %>%
  group_by(loiczid) %>%
  summarize(n_spp = n())

@@

values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
# values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
rast_spp_e  <- crop(rast_spp_e, extent(c(85, 220, -59, 50)))

rast_spp_map_e <- 
  tm_shape(rast_spp_e) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 200, 300, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_CODE_1',
            col = 'orangered4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .35) +
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            frame = FALSE,
            legend.text.size = .35,
            legend.title.size = .45,
            # title.position = 'TOP', 
            legend.outside = TRUE,
            # legend.position = c('right', 'bottom'),
            # legend.bg.color = 'white',
            # legend.bg.alpha = .5,
            attr.outside = TRUE)

@@

save_tmap(rast_spp_map_e, file.path(dir_fig, 'map_am_edge_map.png'), 
          width = 9, height = 6, units = 'cm', dpi = 600)

>>>>

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
                            layer = 'fao_rgns2')

rast_ext  <- extent(c(-30, 330, -75, 75))
fao_polys <- crop(fao_polys, rast_ext)

# wrld_simpl_shift <- wrld_simpl %>%
#   raster::shift(360)
# 
# World1 <- rgeos::gUnion(wrld_simpl, wrld_simpl_shift, byid = FALSE) %>%
#   crop(rast_ext) %>%
#   SpatialPolygonsDataFrame(data.frame(x = 1))
# 
# maptools::writePolyShape(World1, 'clip_fao/fao_rgns/world1')

World1 <- maptools::readShapePoly('clip_fao/fao_rgns/world1')

@@

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum

# spp_cells_clipped <- am_spp_cells %>%
#   filter(am_sid %in% spp_examples_e$am_sid) %>%
#   group_by(loiczid) %>%
#   summarize(n_spp = n())
# write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))

spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))

@@

values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
# values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
rast_spp_e  <- crop(rast_spp_e, rast_ext)

rast_spp_map_e <- 
  tm_shape(rast_spp_e) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 200, 300, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_AREA_1',
            col = 'orangered4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .35) +
  tm_layout(# basemaps = "Esri.WorldTopoMap", 
            frame = TRUE,
            legend.text.size = .35,
            legend.title.size = .45,
            # title.position = 'TOP', 
            legend.outside = FALSE,
            legend.position = c('right', 'top'),
            legend.bg.color = 'white',
            legend.bg.alpha = .8,
            attr.outside = TRUE)

@@

save_tmap(rast_spp_map_e, 
          file.path(dir_fig, 'map_am_edge_map.png'), 
          width = 19, 
          # height = 6, 
          units = 'cm', dpi = 600)

### 2016-09-02 data_explore/data_explore.Rmd

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Paired' = 2.5)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Paired' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(x = 'Class (in order of decreasing number of included species)',
       y = 'Percent of mapped species',
       fill = 'Dataset')

>>>>

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 3)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey70', 'Paired' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(x = 'Class (in order of decreasing number of included species)',
       y = 'Percent of mapped species',
       fill = 'Dataset')

### 2016-09-02 ms_2_methods.Rmd

The IUCN publishes species range maps developed by species experts.  These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A)

@@

@@

For each paired map species, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).  $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.  Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors.

>>>>

The IUCN publishes species range maps developed by species experts.  These experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).  <!--The result is essentially an extent of occurrence (EOO) map for the species-->

@@

<!--In addition to these methodological differences, the two datasets also differ in intent.  IUCN maps essentially communicate the extent of occurrence (EOO) while AquaMaps uses suitable habitat as an indicator of probability of occurrence, -->

@@

For each paired map species, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).  $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.  Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors.  [range area is one critical feature/use of IUCN used as a criterion to help define extinction risk; can AquaMaps provide useful information here as well?]
